P4IRS: A Code Generation Framework for Parallel and Performance-Portable Particle Simulations

### 2025-05-06 sage_latex_template/manuscript.tex
Add figure with the IR example
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>
Add caption
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

@@

\RMchange{Note the domain-specific elements in the IR such as \emph{ParticleInteraction} and \emph{ParticleFor}, as already discussed these can be used during compilation passes for obtaining domain knowledge, and at some point they are lowered according to the strategies specified by the user.
For instance, the \emph{ParticleInteraction} abstract element can be lowered to a code representation that traverses pair of particles by either using the Linked Cells or Verlet Lists data structures.}
%Other elements such as the \emph{ParticleFor} and \emph{ParticleInteraction} are used as higher-level elements of the AST.
%These will be lowered at some point based on their strategies and given parameters.
In this case, the \emph{ParticleInteraction} is lowered by using the linked-cells data structure to build the neighbor-lists as is done in traditional implementations.
The interaction object also contains methods to access the current particle (i), the neighbor particle (j) and the shape for the neighbor particle (shape).
The latter is particulary important for the shape-partitioning and multi-kernel code generation discussed in the \nameref{sec:shape_partitioning} section.

>>>>

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.75\textwidth]{pairs_ir.pdf}
  \caption{\RMchange{P4IRS IR for the BuildNeighborLists example. Red boxes represent elements that are specific to particle simulations, and are used for domain-specific transformations in the \ac{AST}, at some point these elements are lowered to generic, primitive elements that should be straightforward to map into code in a general-purpose language or even an IR such as the LLVM IR. Green boxes represent some of these lower level, generic elements. Blue boxes represent symbols from the simulation, and the dashed lines in red show when these symbols come from a particle-specific element. Elements in yellow represent constants, and both ``s'' elements represent a pattern that is repeated in the constructed \ac{AST} according to the referenced loop in their respective orange boxes.}}
  \label{fig:pairs_ir}
\end{figure*}

@@

\autoref{fig:pairs_ir} depicts a schematic diagram of the constructed P4IRS \ac{IR} for the \emph{BuildNeighborLists} example.
\RMchange{Note the domain-specific elements in the IR such as \emph{ParticleInteraction} and \emph{ParticleFor}, as already discussed these can be used during compilation passes for obtaining domain knowledge, and at some point they are lowered according to the strategies specified by the user.
For instance, the \emph{ParticleInteraction} abstract element can be lowered to a code representation that traverses pair of particles by either using the Linked Cells or Verlet Lists data structures.}
%Other elements such as the \emph{ParticleFor} and \emph{ParticleInteraction} are used as higher-level elements of the AST.
%These will be lowered at some point based on their strategies and given parameters.
In this case, the \emph{ParticleInteraction} is lowered by using the linked-cells data structure to build the neighbor-lists as is done in traditional implementations.
The interaction object also contains methods to access the current particle (i), the neighbor particle (j) and the shape for the neighbor particle (shape).
The latter is particulary important for the shape-partitioning and multi-kernel code generation discussed in the \nameref{sec:shape_partitioning} section.
